---
title: "final project"
author: "Alana Flenoy"
date: "November 10, 2025"
output: slidy_presentation
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)    # for reading CSVs 
library(readxl)   # for reading Excel files 
library(dplyr)    # for data manipulation 
library(ggplot2)  # for making graphs
library(knitr)
library(kableExtra)       # for datatable
```

```{r message = F, echo = F}
chapters <- read_csv("Chapters.csv")  
movies <- read_csv("Movies.csv") 
Spells <- read_csv("Spells.csv")
spells <- Spells %>% select(where(~ !all(is.na(.))))
places <- read_csv("Places.csv") 
Characters <- read.csv("Characters.csv", sep = ";", header = TRUE)
characters <- Characters %>% mutate(across(where(is.character), ~replace_na(.x, "Unknown"))) 
dialogue <- read_excel("Dialogue.xlsx")
Creatures <- read_csv("Creatures.csv")
creatures <- Creatures  %>% select(where(~ !all(is.na(.))))
```

# Data 

```{r}
head(chapters)
head(movies)
head(spells)
head(places)
head(characters)
head(dialogue)
head(creatures)
```

```{r}

```

```{r}

```

# Introduction

Harry Potter is one of the most popular movie franchises of all time. The Harry Potter world 
captivates its audience and leaves them with a sense of wonder and bewilderment. The data being analyzed
will cover the characters, spells, places and dialogue within the books and movies, as well as the 
results of the movies shown through the budget and box office for each movie.


# Questions

  - What percentage of gender, species,  is in each house and school
  - How often does each character's name comes up in the dialogue?
  - What percentage of characters are in each house and/or school?
  - What is the most and least common wand type and how rare is the wand core?
  - How much of each species there is?
  - What are the most common and uncommon patronuses?
  - What are the percentages of blood status in each house and school?

# Key Terms

## Spells

### Charm, Jinx, Spell, or Curse

  - charm: a spell that adds properties to an object or person without changing its fundamental nature
  - jinx: a minor type of dark magic that causes inconvenience or mild harm to the target 
  - spell: a command that uses a specific incantation and wand movement to produce a magical effect     
  - curse: a command that uses a specific incantation and wand movement to produce a magical effect     
  - offensive: are used to attack or harm an opponent 
  - defensive: are used to protect against attacks      

## Creatures

### Mortality

  - extant: still in existence 
  - mortal: able to die 
  - immortal: not able to die 
  - amortal: immune to aging and disease but not to catastrophic physical harm, or a state of not         being alive or        dead 



### Ministry of Magic Classification

  - x creatures: harmless to humans 
  - xx creatures: harmless and could be domesticated 
  - xxx creatures: should be no trouble to competent wizards 
  - xxxx creatures: dangerous and require special knowledge, only trained wizards should handle them,     endangered or just too difficult to capture, rather than being dangerous. 
  - xxxxx creatures: are known wizard killers, are impossible to train or domesticate. 

# House Distribution
## How many characters are in which Hogwarts house? (bar chart)

```{r}

```


# Blood-Status
## What is the percentage of blood statuses? (bar chart)

```{r}

```


# Wands

## What is the least and most common wand wood and core type? (plot)
I need to find a way to separate the wand length, wood, and core.
```{r}
# length wood core
str_split_fixed(characters$Wand, "\"", n = 2)

```

# Places
## What is the top 10 most popular places? (table)
I need to see how often each place is mentioned in the dialogue.
```{r}

```

## Where every place mentioned is found? (bar chart)
Diagon Alley, Dwellings, Hogsmeade, Hogwarts, Other Magic Locations 
```{r}
place_counts <- places %>% 
  count(`Place Category`)

ggplot(place_counts, aes(x = `Place Category`, y = n, fill = `Place Category`)) + 
  geom_bar(stat = "identity") + 
  theme_minimal() + 
  labs( title = "Number of Places per Category", x = "Place Category", y = "Number of Places") 
```

# Spells
## What category do these spells fall into? (scroll box table)
```{r}
type_counts <- spells %>% 
  count(`Type`)

ggplot(type_counts, aes(x = `Type`, y = n, fill = `Type`)) + 
  geom_bar(stat = "identity") + 
  theme_minimal() + 
  labs( title = "Spell Type", x = "Type", y = "Amount of Spell Type") 
```


  - charm: a spell that adds properties to an object or person without changing its fundamental nature
  - jinx: a minor type of dark magic that causes inconvenience or mild harm to the target 
  - spell: a command that uses a specific incantation and wand movement to produce a magical effect  
  - curse: a command that uses a specific incantation and wand movement to produce a magical effect   

---

## Is the spell offensive or defensive? (bar chart)
Add in a column of defensive or offensive
```{r}

```
  - offensive: are used to attack or harm an opponent 
  - defensive: are used to protect against attacks  

---

## Does the color of the light emmitted from the spell that was cast correlate with which type of spell it is?
```{r}
spell_name <- spells$`Spell Name`
spell_incantation <- spells$Incantation
spell_type <- spells$Type
spell_light <- spells$Light
kable(cbind(spell_incantation, spell_type, spell_light), "html") %>%
    kable_styling() %>%
    scroll_box(width = "1000px", height = "500px")
```


# Patronus
## How many characters have a certain patronus? (plot)
```{r}

```


# Creatures
## What creature are the characters classified as? 
```{r}

```

## What is the distribution of mortality and status? (bar)
### Mortality
```{r}
mortality <- creatures %>% 
  count(mortality)

ggplot(mortality, aes(x = `mortality`, y = n, fill = `mortality`)) + 
  geom_bar(stat = "identity") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs( title = "Distribution of Creatures' Mortality", x = "Type of Mortality", y = "Number of Creatures") 
```
  - mortal: able to die 
  - immortal: not able to die 
  - amortal: immune to aging and disease but not to catastrophic physical harm, or a state of not         being alive or        dead 

---

### Status of Existence
```{r}
status <- creatures %>% 
  count(status)

ggplot(status, aes(x = `status`, y = n, fill = `status`)) + 
  geom_bar(stat = "identity") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs( title = "Distribution of Creatures' Status of Existence", x = "Type of Status", y = "Number of Creatures") 
```
  
  - extant: still in existence 
  - endangered: at risk of extinction
  - extinct: does not exist anymore

---

## What is the distribution of the Ministry of Magic classifications? (bar)
```{r}
min_of_mag_class <- creatures %>% 
  count(`ministry of magic classification`)

ggplot(min_of_mag_class, aes(x = `ministry of magic classification`, y = n, fill = `ministry of magic classification`)) + 
  geom_bar(stat = "identity") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs( title = "Creatures' Classification by the Ministry of Magic", x = "Classification", y = "Number of Creatures") 
```
  - beast: classified as a beast but wasn't classified based on how harmful they are
  - x beast: harmless to humans 
  - xx beast: harmless and could be domesticated 
  - xxx beast: should be no trouble to competent wizards 
  - xxxx beast: dangerous and require special knowledge, only trained wizards should handle them,     endangered or just too difficult to capture, rather than being dangerous 
  - xxxxx beast: are known wizard killers, are impossible to train or domesticate 
  - being: any creature that has sufficient intelligence to understand the laws of the magical community and to bear part of  the responsibility in shaping those laws
  - spirit: any creature that has a spectral body, is unaffected by gravity and is not actually alive
  - non-being: are not born and can not be killed

---

## Where are the creatures native to? (bar)
by continent
```{r}

```


# Movies
## What is the correlation between the movie number and run-time? (bar maybe)
```{r}

```


## What is the return of investment? (line table)
return = (future value - initial value) / initial value

```{r}

```

























